### 2023-10-12 cover_letter/cover_letter.tex
Create cover_letter.tex

\documentclass{letter}

\signature{Daiki Tagami}

\address{Big Data Institute\\University of Oxford\\UK}
\begin{document}

\begin{letter}{Bioinformatics}

\opening{Dear Editors,}

I am writing on behalf of my coauthors to submit our manuscript entitled
\emph{tstrait: a quantitative trait simulator for ancestral recombination graphs}
for consideration for publication in Bioinformatics
as an ``Application Note'' research article.

\texttt{tstrait} is a newly developed Python package to simulate quantitative traits from an ancestral recombination graph (ARG). ARG is widely used in population genetics and statistical genetics due to its advantages in genome-wide association study (GWAS), but there does not exist any simulators that can efficiently simulate quantitative traits directly from an ARG. \texttt{tstrait} can simulate quantitative traits at a substantially faster computational speed than traditional techniques, making it to be an essential piece of infrastructure in ARG and GWAS.

We believe that this manuscript provides an overview of current works in ARG and GWAS, and presents the importance of using \texttt{tstrait} in genomic studies. It will be of substantial interest to the readers of Bioinformatics, and we hope that you agree.

\closing{Sincerely,}

\end{letter}
\end{document}